Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-07-07 paper/oopsla.tex
drf hacking
cocktail time. only one todo left, and it is in an appendix

\subsection{Comparison to ``A Promising Semantics 2.1'' [POPL 2017]}
\label{sec:promising}

[Todo.]

Case analysis gives very weak results when combined with thread inlining.
See \cite[\textsection B.1]{DBLP:journals/pacmpl/ChakrabortyV19appendix}.
These happen by performing transformations that: 
(1) introduce conditionals,
(2) inline two threads on both sides of the introduced conditional,
(3) choose different orders for the two threads for the two sides of the conditional.

Case analysis gives very weak results when combined with read introduction.
See \cite{promising-ldrf}.
These happen by performing transformations that: 
(1) introduce reads,
(2) introduce conditionals,
(3) choose different values for the reads on the two sides of the conditional.

The fact that the semantics is not verifiable a posteriori is something it
shares with \weakestmo{}, where the justification relation must be built
inductively.

\weakestmo{} admits FADD, but \PS{} does not.
\PS{} admits CohCYC, but \weakestmo{} does not.

@@

\jjr{} uses substitution rather than Skolemizing.  Indeed our use of
Skolemization is motivated by disjunction closure for predicate transformers,
which do not appear in \jjr{}.  In \reffig{fig:sem}, 
we gave the semantics of read for nonempty pomsets as:
\begin{enumerate}
\item[{\labeltext[\textsc{r}4a]{(\textsc{r}4a)}{read-tau-dep-oopsla}}]
  if $(\aEvs\cap\bEvs)\neq\emptyset$ then
  \begin{math}
    \aTr{\bEvs}{\bForm} \riff
    \aVal{=}\aReg
    \limplies \bForm
  \end{math},    
\item[{\labeltext[\textsc{r}4b]{(\textsc{r}4b)}{read-tau-ind-oopsla}}]
  if $(\aEvs\cap\bEvs)=\emptyset$ then
  \begin{math}
   \aTr{\bEvs}{\bForm} \riff
    \PBR{\aVal{=}\aReg \lor \aLoc{=}\aReg} \limplies
    \bForm.
  \end{math}
\end{enumerate}
In \jjr{}, the definition is roughly as follows:
% (adding the case for $\ref{L6}$, which was missing):
\begin{enumerate}
\item[{\labeltext[\textsc{r}4a$'$]{(\textsc{r}4a$'$)}{read-tau-dep-oopsla-sub}}]
  if $(\aEvs\cap\bEvs)\neq\emptyset$ then
  \begin{math}
    \aTr{\bEvs}{\bForm} \riff
    \bForm[\aVal/\aReg][\aVal/\aLoc]
    % \aVal{=}\aReg
    % \limplies \bForm[\aReg/\aLoc]
  \end{math},    
\item[{\labeltext[\textsc{r}4b$'$]{(\textsc{r}4b$'$)}{read-tau-ind-oopsla-sub}}]
  if $(\aEvs\cap\bEvs)=\emptyset$ then
  \begin{math}
    \aTr{\bEvs}{\bForm} \riff
    \bForm[\aVal/\aReg][\aVal/\aLoc]\land\bForm[\aLoc/\aReg]
  \end{math}
\end{enumerate}
The use of conjunction in \ref{read-tau-ind-oopsla-sub} causes disjunction closure to fail
because the predicate transformer
% $\aTr{}{\bForm}=\bForm[\aVal/\aReg][\aVal/\aLoc]\land\bForm[\aLoc/\aReg]$ does not distribute through
% disjunction:
% \begin{math}
%   \aTr{}{\bForm_1\lor \bForm_2}=
%   (\bForm_1\lor \bForm_2)[\aVal/\aReg][\aVal/\aLoc]\land(\bForm_1\lor \bForm_2)[\aLoc/\aReg]
%   \neq
%   (\bForm_1[\aVal/\aReg][\aVal/\aLoc]\land\bForm_1[\aLoc/\aReg]) \lor
%   (\bForm_2[\aVal/\aReg][\aVal/\aLoc]\land\bForm_2[\aLoc/\aReg])
%   = \aTr{}{\bForm_1} \lor \aTr{}{\bForm_2}
% \end{math}
$\aTr{}{\bForm}=\bForm'\land\bForm''$ does not distribute through
disjunction, even assuming that the prime operations do:\footnote{%
  \begin{math}
    (\bForm_1\lor \bForm_2)'=(\bForm_1'\lor \bForm_2')
  \end{math}
  and
  \begin{math}
    (\bForm_1\lor \bForm_2)''=(\bForm_1''\lor \bForm_2'')
  \end{math}.
}
\begin{math}
  \aTr{}{\bForm_1\lor \bForm_2}=
  \href{https://www.wolframalpha.com/input/?i=\%28a+or+b\%29+and+\%28c+or+d\%29}{(\bForm_1'\lor \bForm_2')\land(\bForm_1''\lor \bForm_2'')}
  \neq
  \href{https://www.wolframalpha.com/input/?i=\%28a+and+c\%29+or+\%28b+and+d\%29}{(\bForm_1'\land\bForm_1'') \lor (\bForm_2'\land\bForm_2'')}
  = \aTr{}{\bForm_1} \lor \aTr{}{\bForm_2}
\end{math}.
% \begin{math}
%   (\bForm_{1}^{1}\lor \bForm_{1}^{2}) \land (\bForm_{2}^{1}\lor \bForm_{2}^{2})
%   \neq
%   (\bForm_{1}^{1}\land\bForm_{2}^{1}) \lor (\bForm_{1}^{2}\land\bForm_{1}^{2}).
% \end{math}
See also \textsection\ref{sec:ex:assoc}.

The substitutions collapse $\aLoc$ and $\aReg$, allowing local invariant
reasoning (\xLIR{}), as required by causality test case 1, discussed at the end of
\textsection\ref{sec:ex:control}.  Without Skolemizing it is necessary to
substitute $[\aLoc/\aReg]$, since the reverse substitution $[\aReg/\aLoc]$ is
useless when $\aReg$ is bound---compare with
\textsection\ref{sec:substitutions}.  As discussed below (\ref{p:downset}),
including this substitution affects the interaction of \xLIR{} and downset
closure.

>>>>

\subsection{Comparison to ``Promising Semantics'' [POPL 2017]}
\label{sec:promising}

\citet{DBLP:journals/pacmpl/JagadeesanJR20} note that the promising semantics
(\PS{}) \cite{DBLP:conf/popl/KangHLVD17} and related models
\citep{DBLP:conf/esop/JagadeesanPR10,DBLP:journals/pacmpl/ChakrabortyV19,Manson:2005:JMM:1047659.1040336},
fail to validate compositional reasoning of temporal properties.  Consider
their example \ref{oota4}:
\begin{align}
  \taglabel{oota4}
  \begin{gathered}
    \PW{y}{x}
    \PAR
    \PR{y}{r} \SEMI \IF{b}\THEN  \PW{x}{r} \SEMI \PW{z}{r} \ELSE \PW{x}{1} \FI
    \PAR
    \PW{b}{1}
    \\[-1ex]
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{rx}{\DR{x}{1}}{}
        \event{wy}{\DW{y}{1}}{right=of rx}
        \po{rx}{wy}
        \event{ry}{\DR{y}{1}}{right=3em of wy} 
        \event{wx}{\DW{x}{1}}{right=of ry}
        \event{wz}{\DW{z}{1}}{right=of wx}
        \event{rb}{\DR{b}{1}}{right=of wz}
        \event{wb1}{\DW{b}{1}}{right=3em of rb}
        \po{ry}{wx}
        \rf{wb1}{rb}
        \rf{wy}{ry}
        \rf[out=-170,in=-10]{wx}{rx}
        \po{rb}{wz}
        \po[out=15,in=165]{ry}{wz}
      \end{tikzinline}}
  \end{gathered}  
\end{align}
Under \PS{}, this outcome is allowed by \emph{baiting} with the \texttt{else}
branch, then \emph{switching} to the \texttt{then} branch, based on a coin
flip $(\IF{b})$.\footnote{Call the threads \texttt{s}, \texttt{t}, and
  \texttt{u}.  To get the result in the promising semantics, first execute
  \texttt{u} to get message \texttt{<b:1@1>}.  Then \texttt{t} promises
  \texttt{<x:1@1>}, which it can fulfill by reading \texttt{b}$=$\texttt{0}.
  Then execute \texttt{s} to get message \texttt{<y:1@1>}.  Then execute
  \texttt{t}, reading \texttt{b}$=$\texttt{1} and \texttt{y}$=$\texttt{1} and
  fulfill the promise by writing \texttt{<x:1@1>}. The execution is exactly
  the same in our speculative semantics \cite{DBLP:conf/esop/JagadeesanPR10},
  removing timestamps and replacing the word \emph{promise} by
  \emph{speculation}.}

Under all variants of \PwT{}, this outcome is disallowed, due to a cycle in
${\rrfx}\cup{\lt}$. (Note that all reads cross threads in the example, so
there is no difference between \PwTmca{1} and \PwTmca{2}.)

\citet[Fig.~8]{DBLP:journals/toplas/Lochbihler13} notes that such violations
of temporal reasoning could result in violations of the Java security
architecture.

The difference between \PS{} and \PwT{} can be understood in terms of the
valid program transformations.  \PS{} allows reads to be introduced, with
subsequent case analysis (aka, if-closure) on the value read---effectively,
this can turn one read into two, with different conditional branches taken
for the two copies of the read.  For detailed examples, see
\cite{promising-ldrf}.

\PwT{} invalidates read introduction.  In return, \PwT{} enjoys
compositionality for temporal safety properties, as discussed in
\textsection\ref{sec:results} and
\cite[\textsection6]{DBLP:journals/pacmpl/JagadeesanJR20}.

This difference highlights the subtle tensions in relaxed memory models.
It is not possible to have everything one wants, thus, one is forced
to choose which optimizations and reasoning principles are most important.

% \todo{Write this.}

% Case analysis gives very weak results when combined with thread inlining.
% See \cite[\textsection B.1]{DBLP:journals/pacmpl/ChakrabortyV19appendix}.
% These happen by performing transformations that: 
% (1) introduce conditionals,
% (2) inline two threads on both sides of the introduced conditional,
% (3) choose different orders for the two threads for the two sides of the conditional.

% Case analysis gives very weak results when combined with read introduction.
% See \cite{promising-ldrf}.
% These happen by performing transformations that: 
% (1) introduce reads,
% (2) introduce conditionals,
% (3) choose different values for the reads on the two sides of the conditional.

% The fact that the semantics is not verifiable a posteriori is something it
% shares with \weakestmo{}, where the justification relation must be built
% inductively.

% \weakestmo{} admits FADD, but \PS{} does not.
% \PS{} admits CohCYC, but \weakestmo{} does not.

@@

\jjr{} uses substitution rather than Skolemizing.  Indeed our use of
Skolemization is motivated by disjunction closure for predicate transformers,
which do not appear in \jjr{}.  In \reffig{fig:seq}, 
we gave the semantics of read for nonempty pomsets as:
\begin{enumerate}
\item[{\labeltext[\textsc{r}4a]{(\textsc{r}4a)}{read-tau-dep-oopsla}}]
  if $(\aEvs\cap\bEvs)\neq\emptyset$ then
  \begin{math}
    \aTr{\bEvs}{\bForm} \riff
    \aVal{=}\aReg
    \limplies \bForm
  \end{math},    
\item[{\labeltext[\textsc{r}4b]{(\textsc{r}4b)}{read-tau-ind-oopsla}}]
  if $(\aEvs\cap\bEvs)=\emptyset$ then
  \begin{math}
   \aTr{\bEvs}{\bForm} \riff
    \PBR{\aVal{=}\aReg \lor \aLoc{=}\aReg} \limplies
    \bForm.
  \end{math}
\end{enumerate}
In \jjr{}, the definition is roughly as follows:
% (adding the case for $\ref{L6}$, which was missing):
\begin{enumerate}
\item[{\labeltext[\textsc{r}4a$'$]{(\textsc{r}4a$'$)}{read-tau-dep-oopsla-sub}}]
  if $(\aEvs\cap\bEvs)\neq\emptyset$ then
  \begin{math}
    \aTr{\bEvs}{\bForm} \riff
    \bForm[\aVal/\aReg][\aVal/\aLoc]
    % \aVal{=}\aReg
    % \limplies \bForm[\aReg/\aLoc]
  \end{math},    
\item[{\labeltext[\textsc{r}4b$'$]{(\textsc{r}4b$'$)}{read-tau-ind-oopsla-sub}}]
  if $(\aEvs\cap\bEvs)=\emptyset$ then
  \begin{math}
    \aTr{\bEvs}{\bForm} \riff
    \bForm[\aVal/\aReg][\aVal/\aLoc]\land\bForm[\aLoc/\aReg]
  \end{math}
\end{enumerate}
The use of conjunction in \ref{read-tau-ind-oopsla-sub} causes disjunction closure to fail
because the predicate transformer
% $\aTr{}{\bForm}=\bForm[\aVal/\aReg][\aVal/\aLoc]\land\bForm[\aLoc/\aReg]$ does not distribute through
% disjunction:
% \begin{math}
%   \aTr{}{\bForm_1\lor \bForm_2}=
%   (\bForm_1\lor \bForm_2)[\aVal/\aReg][\aVal/\aLoc]\land(\bForm_1\lor \bForm_2)[\aLoc/\aReg]
%   \neq
%   (\bForm_1[\aVal/\aReg][\aVal/\aLoc]\land\bForm_1[\aLoc/\aReg]) \lor
%   (\bForm_2[\aVal/\aReg][\aVal/\aLoc]\land\bForm_2[\aLoc/\aReg])
%   = \aTr{}{\bForm_1} \lor \aTr{}{\bForm_2}
% \end{math}
$\aTr{}{\bForm}=\bForm'\land\bForm''$ does not distribute through
disjunction, even assuming that the prime operations do:\footnote{%
  \begin{math}
    (\bForm_1\lor \bForm_2)'=(\bForm_1'\lor \bForm_2')
  \end{math}
  and
  \begin{math}
    (\bForm_1\lor \bForm_2)''=(\bForm_1''\lor \bForm_2'')
  \end{math}.
}
\begin{math}
  \aTr{}{\bForm_1\lor \bForm_2}=
  \href{https://www.wolframalpha.com/input/?i=\%28a+or+b\%29+and+\%28c+or+d\%29}{(\bForm_1'\lor \bForm_2')\land(\bForm_1''\lor \bForm_2'')}
  \neq
  \href{https://www.wolframalpha.com/input/?i=\%28a+and+c\%29+or+\%28b+and+d\%29}{(\bForm_1'\land\bForm_1'') \lor (\bForm_2'\land\bForm_2'')}
  = \aTr{}{\bForm_1} \lor \aTr{}{\bForm_2}
\end{math}.
% \begin{math}
%   (\bForm_{1}^{1}\lor \bForm_{1}^{2}) \land (\bForm_{2}^{1}\lor \bForm_{2}^{2})
%   \neq
%   (\bForm_{1}^{1}\land\bForm_{2}^{1}) \lor (\bForm_{1}^{2}\land\bForm_{1}^{2}).
% \end{math}
See also \textsection\ref{sec:ex:assoc}.

The substitutions collapse $\aLoc$ and $\aReg$, allowing local invariant
reasoning (\xLIR{}), as required by \jmm{} causality test case 1, discussed at the end of
\textsection\ref{sec:ex:control}.  Without Skolemizing it is necessary to
substitute $[\aLoc/\aReg]$, since the reverse substitution $[\aReg/\aLoc]$ is
useless when $\aReg$ is bound---compare with
\textsection\ref{sec:substitutions}.  As discussed below (\ref{p:downset}),
including this substitution affects the interaction of \xLIR{} and downset
closure.

### 2021-07-07 paper/outro.tex
a stab at DRF
cocktail time. only one todo left, and it is in an appendix

\section{Future Work}

@@

In \S\ref{sec:arm1} we presented a compilation strategy to \armeight{}
for a simplified model, but which introduces fences to acquiring
reads. These fences are not required in \S\ref{sec:arm2}, but at the
cost of model complexity. It would be illuminating to find out what
the performance penalty is for these fences.

The promising semantics (\PS{}) validates read introduction whereas \PwT{}
does not.  As a result \PS{} admits behaviors that break reasoning about
temporal safety properties (see \textsection\ref{sec:promising}).
Nonetheless, read introduction is ubiquitous is some compilers.  It would be
interesting to know if there is a performance penalty for banning read
introduction.

An earlier version of this paper has been mechanized in Agda; it would be
reassuring to update the mechanization to bring it in line with the current state.

We don't handle access elimination.

>>>>

\section{Conclusions} %Future Work}

@@

We've also not handled access elimination: store-forwarding and
dead-write-removal are unsound.  We expect that these can be validated by
allowing events with different labels to merge in some instances.

\PwTmca{1} is a simpler model than \PwTmca{2}, but requires fences on
acquiring reads for \armeight{}.  It would be illuminating to find out what
the performance penalty is for these fences.

In a similar vein, the promising semantics (\PS{}) validates read
introduction whereas \PwTmca{} does not.  As a result \PS{} admits behaviors
that break reasoning about temporal safety properties (see
\textsection\ref{sec:promising}).  Nonetheless, read introduction is
ubiquitous is some compilers.  It would be interesting to know if there is a
performance penalty for banning read introduction.

% An earlier version of this paper has been mechanized in Agda; it would be
% reassuring to update the mechanization to bring it in line with the current state.

### 2021-07-07 paper/results.tex
edits
cocktail time. only one todo left, and it is in an appendix

\section{Results}
\label{sec:results}

% \subsection{Reordering Transformations}
% \label{sec:ex:valid}

% When $\sem{\aCmd} \supseteq \sem{\aCmd'}$, we say that $\aCmd'$ is a
% \emph{valid transformation} of $\aCmd$.
% In this subsection, we show the
% validity of specific optimizations.

The semantics validates many peephole optimizations.  Most apply only to
relaxed access.
\begin{align*}
  %\taglabel{RR}
  \sembase{\PR{\aLoc}{\aReg} \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg}\SEMI \PR{\aLoc}{\aReg}} &&\text{if } \aReg\neq\bReg
  \\
  %\taglabel{WW}
  \sembase{\aLoc \GETS \aExp \SEMI \bLoc  \GETS \bExp} &=
  \sembase{\bLoc  \GETS \bExp\SEMI \aLoc \GETS \aExp} &&\text{if } \aLoc\neq\bLoc
  \\
  %\taglabel{RW}
  \sembase{\aLoc \GETS \aExp  \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg} \SEMI\aLoc \GETS \aExp} &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
\end{align*}
% \ref{WW}, \ref{RW} and \ref{RR} require that two sides of the semicolon
% have disjoint ids; for example, \ref{RW} requires $\disjoint{{\free(\aReg
% \GETS \aLoc)}}{{\free(\bLoc \GETS \bExp)}}$. 
% \ref{RR} requires either $\aReg\neq\bReg$ or
% $\aLoc=\bLoc$.  \ref{WW} and \ref{RW} require that two sides of the
% semicolon have disjoint ids; for example, \ref{RW} requires
% $\disjoint{{\free(\PR{\aLoc}{\aReg})}}{{\free(\bLoc \GETS \bExp)}}$.
%\ref{5} imposes no order between events in \ref{RR}--\ref{RW}.  %Note that \ref{RR} allows aliasing.
Here $\free(\aCmd)$ is the set of locations and registers that occur in $\aCmd$.
Using augmentation closure, the semantics also validates roach-motel reorderings \cite{SevcikThesis}.  For
example, on read/write pairs:
  \begin{align*}
    %\tag{\textsc{roach1}}\label{AcqW}
    \sembase{x^\amode \GETS \aExp \SEMI\PR{y}{\bReg}} &\supseteq
    \sembase{\PR{y}{\bReg}  \SEMI x^\amode\GETS \aExp} 
    &&\text{if }
    \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
    \\
    %\tag{\textsc{roach2}}\label{RelW}
    \sembase{x \GETS \aExp \SEMI\PR[\amode]{y}{\bReg}} &\supseteq
    \sembase{\PR[\amode]{y}{\bReg}  \SEMI x\GETS \aExp} 
    &&\text{if }
    \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
  \end{align*}

>>>>

\section{\PwTmcaTITLE{} Results}
\label{sec:results}

\PwP{} \cite{DBLP:journals/pacmpl/JagadeesanJR20} is a novel memory model,
intended to serve as a semantic basis for a Java-like language, where all
access is safe.  \PwTmca{} generalizes \PwP{}, making several small but
significant changes.  As a result, we have had to re-prove most of the
theorems from \PwP{}.

In \textsection\ref{sec:arm}, we show that \PwTmca{1} supports the optimal
lowering of relaxed accesses to \armeight{} and that \PwTmca{2} supports the
optimal lowering of \emph{all} accesses to \armeight{}.  The proofs are based
on two recent characterizations of \armeight{} \cite{armed}.  For \PwTmca{1},
we use \emph{External Global Consistency}.  For \PwTmca{2}, we use
\emph{External Consistency}.

In \textsection\ref{sec:sc}, we prove sequential consistency for
local-data-race-free programs.  The proof uses \emph{program order}, which we
construct for \cXI{} in \textsection\ref{sec:c11}.  The same construction
works for \PwTmca{}.  (This proof assumes there are no \RMW{} operations.)

The semantics validates many peephole optimizations, such as the standard reorderings on relaxed access:
\begin{align*}
  %\taglabel{RR}
  \sembase{\PR{\aLoc}{\aReg} \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg}\SEMI \PR{\aLoc}{\aReg}} &&\text{if } \aReg\neq\bReg
  \\
  %\taglabel{WW}
  \sembase{\aLoc \GETS \aExp \SEMI \bLoc  \GETS \bExp} &=
  \sembase{\bLoc  \GETS \bExp\SEMI \aLoc \GETS \aExp} &&\text{if } \aLoc\neq\bLoc
  \\
  %\taglabel{RW}
  \sembase{\aLoc \GETS \aExp  \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg} \SEMI\aLoc \GETS \aExp} &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
\end{align*}
% \ref{WW}, \ref{RW} and \ref{RR} require that two sides of the semicolon
% have disjoint ids; for example, \ref{RW} requires $\disjoint{{\free(\aReg
% \GETS \aLoc)}}{{\free(\bLoc \GETS \bExp)}}$. 
% \ref{RR} requires either $\aReg\neq\bReg$ or
% $\aLoc=\bLoc$.  \ref{WW} and \ref{RW} require that two sides of the
% semicolon have disjoint ids; for example, \ref{RW} requires
% $\disjoint{{\free(\PR{\aLoc}{\aReg})}}{{\free(\bLoc \GETS \bExp)}}$.
%\ref{5} imposes no order between events in \ref{RR}--\ref{RW}.  %Note that \ref{RR} allows aliasing.
Here $\free(\aCmd)$ is the set of locations and registers that occur in $\aCmd$.
Using augmentation closure, the semantics also validates roach-motel reorderings \cite{SevcikThesis}.  For
example, on read/write pairs:
\begin{align*}
  % \tag{\textsc{roach1}}\label{AcqW}
  \sembase{x^\amode \GETS \aExp \SEMI\PR{y}{\bReg}} &\supseteq
  \sembase{\PR{y}{\bReg}  \SEMI x^\amode\GETS \aExp} 
  &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
  \\
  % \tag{\textsc{roach2}}\label{RelW}
  \sembase{x \GETS \aExp \SEMI\PR[\amode]{y}{\bReg}} &\supseteq
  \sembase{\PR[\amode]{y}{\bReg}  \SEMI x\GETS \aExp} 
  &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
\end{align*}

Notably, the semantics does \emph{not} validate read introduction.  When
combined with case analysis (\textsection\ref{sec:semca}), read introduction
can break temporal reasoning.  This combination is allowed by speculative
operational models.  See \textsection\ref{sec:promising} for a discussion.

Prop.~6.1 of \cite{DBLP:journals/pacmpl/JagadeesanJR20} establishes a
compositional principle for proving that programs validate formula in
past-time temporal logic.  The principal is based entirely on the pomset
order relation.  It's proof, and all of the no-thin-air examples in
\cite[\textsection6]{DBLP:journals/pacmpl/JagadeesanJR20} hold equally for
the models described here.
